---
title: "Final Project"
date: "2019/11/10"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
library(tidyverse)
library(plotly)
library(rvest)
library(readr)
library(viridis)
library(leaflet)
library(patchwork)
library(RColorBrewer)
library(corrplot)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom",plot.title = element_text(face = "bold")))
```


## load collision data
```{r}
data_2018 = 
  read_csv("./data/2018data.csv") %>% 
  janitor::clean_names() 
newnames = colnames(data_2018) %>% 
  str_replace("number_of_","") 
names(data_2018) = newnames

tidy_data = 
data_2018 %>% 
  mutate(
    date_complete = date 
  ) %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  separate(time, into = c("hour", "minute"), sep = ":") %>%
  mutate(
    date_paste = as.Date(paste("2018",month,day,sep = "-"))
  ) %>% 
  select(-zip_code, -location, -on_street_name, -cross_street_name, -off_street_name,-collision_id,-year) %>% 
  rename("vehicle_type" = "vehicle_type_code_1",
         "cause" = "contributing_factor_vehicle_1") %>% 
  mutate( day = as.numeric(day),
          month = as.numeric(month),
          hour = as.numeric(hour),
          minute = as.numeric(minute),
          latitude = replace_na(latitude,0),
          vehicle_type = str_to_lower(vehicle_type)
  ) %>%
  filter( latitude != 0) %>% 
  drop_na(borough)
  
  
```


### Vehicle type
```{r}
vehicle_type_data = 
  tidy_data %>% 
  mutate(
    vehicle_type = replace(vehicle_type,str_detect(vehicle_type,"truck"),"truck"),
    vehicle_type = replace(vehicle_type,str_detect(vehicle_type,"sport utility"),"sport utility vehicle"),
    vehicle_type = replace(vehicle_type, vehicle_type %in% c("taxi","passenger vehicle","sedan","truck","sport utility vehicle") == FALSE, "others")) %>% 
  group_by(vehicle_type,hour) %>% 
  summarize(
    n = n()
  )

vehicle_type_plot = 
vehicle_type_data %>% 
  ggplot(aes(x = hour, y = n, color = vehicle_type)) +
  geom_line() +
  labs(
    color = "vehicle type",
    title = "Collisions of Day for Different Vehicles",
    x = list(title = "Hour of Day"),
    y = list(title = "Collisions")
    )
  
ggplotly(vehicle_type_plot)
```


### Top 10 Collision Reasons
```{r}
# reason_data = 
#   tidy_data %>%
#   group_by(contributing_factor_vehicle_1) %>%
#   summarize(n = n()) %>% 
#   arrange(desc(n)) %>% 
#   head(10)
# reason_data %>% 
#   plot_ly(x = ~reorder(contributing_factor_vehicle_1,desc(n)), y = ~n, color = ~contributing_factor_vehicle_1 ,type = "bar") %>% 
#   layout(
#     title = "The Number of Items Ordered in Each Aisle",
#     xaxis = list(title = "Different Reasons"),
#     yaxis = list(title = "Count")
#     )

df_cause =  
  tidy_data %>% 
  select(borough,cause) %>% 
      filter(!(cause %in% c('', 
                            'Unspecified', 
                            'Driver Inattention/Distraction',
                            'Other Vehicular',
                            'Failure to Yield Right-of-Way'))
             ) 
df_cause_top = 
  df_cause %>%
  group_by(cause) %>%
  summarise(count = n()) %>%
  top_n(10, count) %>% 
  ungroup() %>% 
  arrange(count)

cause_by_boro =  
  df_cause %>% 
  semi_join(., df_cause_top, by = "cause") %>%
  group_by(borough, cause) %>%
  summarise(count = n()) %>% 
  # reorder cause levels for desirable order in bar chart display:
  mutate(
    cause = factor(cause, levels = pull(df_cause_top,cause))
    ) 

cause_by_boro_all = 
  cause_by_boro %>% 
  group_by(borough) %>%
  summarise(total = sum(count))

cause_by_boro = 
  left_join(cause_by_boro, cause_by_boro_all, by = "borough") %>%
  mutate(ratio = count / total) %>%
  select(-count, -total) 

cause_by_boro_plot =
  cause_by_boro %>% 
  ggplot(aes(x = cause, y = ratio, fill = borough)) +
      geom_bar(width = 0.5, stat = 'identity', show.legend = F) +
      coord_flip() +
      facet_grid(.~borough) +
      labs(title = 'Collisions by Causes',
           x = NULL,
           y = NULL) +
  theme(axis.text.x = element_text(angle = 45))
ggplotly(cause_by_boro_plot)
```

## Mapping

```{r}
data_2018 = tidy_data
map_data = rename(data_2018, long = latitude, lat = longitude) 
pal <- colorNumeric(
  palette = "Accent",
  domain = map_data$persons_injured)
map_data %>% 
  filter(persons_injured > 0) %>% 
  filter(!(lat < "-70" | lat >= "-75")) %>% 
  mutate(
    label = str_c("<b>vehicle type: ", vehicle_type, "</b><br>Persons Injured: ", persons_injured , sep = "") ) %>%
  leaflet() %>% 
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addLegend("bottomright", pal = pal, values = ~persons_injured,
    title = "Persons Injured",
    opacity = 1
  ) %>% 
  addCircleMarkers(
    ~lat, ~long,
    color = ~pal(persons_injured),
    radius = 0.5,
    popup = ~ label) 
```

```{r}
data_2018_seperate = tidy_data 
```



### Association between Day and People Injured or killed or Accident Number

```{r, echo = FALSE}
data_acci_day = tidy_data %>% 
  group_by(date_paste) %>% 
  mutate(
    sum_acci = n()
  ) %>% 
  select(date_paste, sum_acci) %>% 
  distinct()

data_kill_injured_day = tidy_data %>% 
  group_by(date_paste) %>% 
  mutate(
    sum_injured = sum(persons_injured,na.rm = TRUE),
    sum_killed = sum(persons_killed,na.rm = TRUE)
  ) %>% 
  select(date_paste, sum_injured, sum_killed) %>% 
  distinct( ) 

data_day = cbind(data_acci_day, data_kill_injured_day) %>% 
  select(-date_paste1)
data_day= rename(data_day, 
                "accident number"  = "sum_acci",
                "people injured"  = "sum_injured" , 
                 "people killed" = "sum_killed" ,
                 "date" = "date_paste" ) 

data_day  = data_day %>% 
  pivot_longer(
    `accident number`:`people killed`,
    names_to   = "type",
    values_to = "number"
  )
```


```{r}
data_day %>%
  filter(type == "accident number") %>% 
  arrange(desc(number))

plot_day =
  data_day %>% 
  ggplot(aes(x = date,y = number, color = type))+
  geom_line()+
  theme(
        axis.title = element_text(size=14,face="bold"),
        plot.title = element_text(hjust = 0.5,size=14,face="bold"))+
  labs(
   title = "Trend Over the Year",
   x = "Day of the Year",
   y = "Number") +
  annotate("text", x = as.Date("2018-11-15") , y = 695, label = "2018-11-15")

plot_kill_injured_day = ggplotly(plot_day)
plot_kill_injured_day
```



### proportion of accident and injured people by borough, by hour

```{r}
boro_events = 
  tidy_data %>% 
  drop_na(borough, persons_injured, persons_killed) %>%
  group_by(borough) %>%
  summarize(accident = n(),
            injured = sum(persons_injured),
            killed = sum(persons_killed)) %>%
  mutate(prop_accident = round(accident/sum(accident),4),
         prop_injured = round(injured/sum(injured),4),
         prop_killed = round(killed/sum(killed),4)) %>%
  pivot_longer(
    prop_accident:prop_killed,
    values_to = "proportion",
    names_prefix = "prop_",
    names_to = "events",
  ) 

boro_events_plot = 
  boro_events %>% 
  ggplot(aes(x = borough, y = proportion, fill = events)) +
  geom_bar(stat = "identity",position = "dodge") +
  labs(title ="proportion of accident and injured people by borough", 
       x = 'Borough',
       y = 'Proportion'
  )

ggplotly(boro_events_plot)

```
Number of accident and injured people in Brooklyn is greatest comparing with other boroughs. Queen is second. Staten Island have fewest accident and injured people. Manhattan and Bronx is in the middle.

```{r}
hour_events = 
  tidy_data %>% 
  drop_na(hour, persons_injured, persons_killed) %>%
  group_by(hour) %>%
  summarize(accident = n(),
            injured = sum(persons_injured),
            killed = sum(persons_killed)) %>%
  mutate(prop_accident = round(accident/sum(accident),4),
         prop_injured = round(injured/sum(injured),4),
         prop_killed = round(killed/sum(killed),4)) %>%
  pivot_longer(
    prop_accident:prop_killed,
    values_to = "proportion",
    names_prefix = "prop_",
    names_to = "events",
  ) 

hour_events_plot =
hour_events %>% 
  ggplot(aes(x = hour, y = proportion, color = events)) +
  geom_line() +
  annotate("rect", xmin = 8, xmax = 20, ymin = 0, ymax = Inf, fill = "blue", alpha = .1) +
  labs(
    title = "proportion of accident and injured people by hour", 
    x =  'Hour',
    y = 'Proportion')

hour_events_plot = ggplotly(hour_events_plot)
hour_events_plot[['x']][['layout']][['shapes']] <- c()

hour_events_plot = layout(hour_events_plot, shapes = list(type = "rect", fillcolor = "pink", line = list(color = "pink"), opacity = 0.3, x0 = 8, x1 = 20, xref = "x", y0 = 0, y1 = 0.08, yref = "y"))

hour_events_plot

```
Frequency of accident and injured people in the period of 8 am-20 pm is higher than other period of time. Started from 3 am, number of accident and injured people increase and reach a small peak at 8am. At 5pm, it reaches a big peak. then it started to decrease.

## Holiday Analysis
```{r}
data_2018 = tidy_data

holiday_data = 
  data_2018 %>%
  select(date_complete, everything()) %>%
  mutate(holiday = "No Holiday") %>% 
  select(date_complete, holiday, everything()) %>% 
  mutate(holiday = replace(holiday, date_complete == "01/01/2018", "New Year's Day"), 
         holiday = replace(holiday, date_complete == "01/15/2018", "Martin Luther King, Jr. Day"),
         holiday = replace(holiday, date_complete == "01/13/2018"|date_complete =="01/14/2018", "Martin Luther King Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "02/19/2018", "Presidents Day"),
         holiday = replace(holiday, date_complete == "02/17/2018"|date_complete =="02/18/2018", "Presidents Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "05/28/2018", "Memorial Day"),
         holiday = replace(holiday, date_complete == "05/26/2018"|date_complete =="05/27/2018", "Memorial Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "07/04/2018", "Independence Day"),
         holiday = replace(holiday, date_complete == "07/07/2018"|date_complete =="07/08/2018", "Independence Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "09/03/2018", "Labor Day"),
         holiday = replace(holiday, date_complete == "09/01/2018"|date_complete =="09/02/2018", "Labor Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "11/22/2018", "Thanksgiving Day"),
         holiday = replace(holiday, date_complete == "11/23/2018"|date_complete =="11/24/2018"|date_complete =="11/25/2018", "Thanksgiving Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "12/25/2018", "Christmas Day"),
         holiday = replace(holiday, date_complete == "12/26/2018"|date_complete =="12/27/2018", "Christmas Day Weekend (2 days)")
         )
```

```{r}
holiday_data =
  holiday_data %>% 
  drop_na(persons_injured)

holiday_data %>% 
  group_by(holiday) %>% 
  summarise(total_injured = sum(persons_injured))

bronx = 
  holiday_data %>% 
  filter(borough == "BRONX") %>% 
  group_by(holiday) %>% 
  summarise(bronx_total_injured = sum(persons_injured))
brooklyn = 
  holiday_data %>% 
  filter(borough == "BROOKLYN") %>% 
  group_by(holiday) %>% 
  summarise(brooklyn_total_injured = sum(persons_injured)) %>% 
  select(brooklyn_total_injured)
manhattan = 
  holiday_data %>% 
  filter(borough == "MANHATTAN") %>% 
  group_by(holiday) %>% 
  summarise(manhattan_total_injured = sum(persons_injured))%>% 
  select(manhattan_total_injured)
queens = 
  holiday_data %>% 
  filter(borough == "QUEENS") %>% 
  group_by(holiday) %>% 
  summarise(queens_total_injured = sum(persons_injured))%>% 
  select(queens_total_injured)
staten_island = 
  holiday_data %>% 
  filter(borough == "STATEN ISLAND") %>% 
  group_by(holiday) %>% 
  summarise(staten_island_total_injured = sum(persons_injured))%>% 
  select(staten_island_total_injured)

holiday_borough = 
  cbind(bronx, brooklyn, manhattan, queens, staten_island) %>% 
  mutate(total = bronx_total_injured+brooklyn_total_injured+manhattan_total_injured+	queens_total_injured+staten_island_total_injured) %>% 
  filter(holiday!="No Holiday")

knitr::kable(holiday_borough)

holiday_borough_day = 
  holiday_borough %>% 
  filter(holiday =="Christmas Day"|holiday =="Independence Day"|holiday =="Labor Day"|holiday =="Martin Luther King, Jr. Day"|holiday =="Memorial Day"|holiday =="New Year’s Day"|holiday =="Thanksgiving Day"|holiday =="Presidents Day")

knitr::kable(holiday_borough_day)

holiday_borough_day = holiday_borough_day[order(holiday_borough_day$total),]

# On the day of the holiday

plot_day =
pivot_longer(
    holiday_borough_day, 
    bronx_total_injured:staten_island_total_injured,
    names_to = "borough", 
    values_to = "person_injured") %>% 
  ggplot(aes(x = reorder(holiday, -total), y = person_injured, fill = borough)) +
  geom_bar(stat="identity") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 45)) +
  labs(
    title = "Persons Injured by Car Crashs \n on Federal Holidays (2018)",
    x = "Holiday",
    y = "# Persons Injured") +
  scale_fill_discrete(name = "Borough", labels = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"))

# scale_fill_discrete(name = "Borough", labels = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island")) 

```

# On three day weekends 

```{r}
data_2018 = tidy_data

holiday_data_weekend = 
  data_2018 %>%
  select(date_complete, everything()) %>%
  mutate(holiday = "No Holiday") %>% 
  select(date_complete, holiday, everything()) %>% 
  mutate(holiday = replace(holiday, date_complete == "01/15/2018"|date_complete == "01/13/2018"|date_complete =="01/14/2018", "Martin Luther King Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "02/17/2018"|date_complete == "02/19/2018"|date_complete =="02/18/2018", "Presidents Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "05/28/2018"|date_complete == "05/26/2018"|date_complete =="05/27/2018", "Memorial Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "07/04/2018"|date_complete == "07/07/2018"|date_complete =="07/08/2018", "Independence Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "09/03/2018"|date_complete == "09/01/2018"|date_complete =="09/02/2018", "Labor Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "11/22/2018"|date_complete == "11/23/2018"|date_complete =="11/24/2018"|date_complete =="11/25/2018", "Thanksgiving Weekend (4 days)"),
         holiday = replace(holiday, date_complete == "12/25/2018"|date_complete == "12/26/2018"|date_complete =="12/27/2018", "Christmas Day Weekend (3 days)")
         )

holiday_data_weekend =
  holiday_data_weekend %>% 
  drop_na(persons_injured)

holiday_data_weekend %>% 
  group_by(holiday) %>% 
  summarise(total_injured = sum(persons_injured))

bronx = 
  holiday_data_weekend %>% 
  filter(borough == "BRONX") %>% 
  group_by(holiday) %>% 
  summarise(bronx_total_injured = sum(persons_injured))
brooklyn = 
  holiday_data_weekend %>% 
  filter(borough == "BROOKLYN") %>% 
  group_by(holiday) %>% 
  summarise(brooklyn_total_injured = sum(persons_injured)) %>% 
  select(brooklyn_total_injured)
manhattan = 
  holiday_data_weekend %>% 
  filter(borough == "MANHATTAN") %>% 
  group_by(holiday) %>% 
  summarise(manhattan_total_injured = sum(persons_injured))%>% 
  select(manhattan_total_injured)
queens = 
  holiday_data_weekend %>% 
  filter(borough == "QUEENS") %>% 
  group_by(holiday) %>% 
  summarise(queens_total_injured = sum(persons_injured))%>% 
  select(queens_total_injured)
staten_island = 
  holiday_data_weekend %>% 
  filter(borough == "STATEN ISLAND") %>% 
  group_by(holiday) %>% 
  summarise(staten_island_total_injured = sum(persons_injured))%>% 
  select(staten_island_total_injured)

holiday_borough_weekend = 
  cbind(bronx, brooklyn, manhattan, queens, staten_island) %>% 
  mutate(total = bronx_total_injured+brooklyn_total_injured+manhattan_total_injured+	queens_total_injured+staten_island_total_injured) %>% 
  filter(holiday!="No Holiday")

knitr::kable(holiday_borough_weekend)

plot_weekend = 
pivot_longer(
    holiday_borough_weekend, 
    bronx_total_injured:staten_island_total_injured,
    names_to = "borough", 
    values_to = "person_injured") %>% 
  ggplot(aes(x = reorder(holiday, -total), y = person_injured, fill = borough)) +
  geom_bar(stat="identity") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 45)) +
  labs(
    title = "Persons Injured by Car Crashs on \n Federal Holidays Weekends (2018)",
    x = "Holiday Weekends",
    y = "# Persons Injured") +
  scale_fill_discrete(name = "Borough", labels = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"))

plot_day + plot_weekend
```


# Weather

```{r}
climate_df = 
  read_csv("./large_data/climate.csv") %>% 
  janitor::clean_names() %>% 
  select(date,hourly_dry_bulb_temperature,hourly_precipitation,daily_weather) %>%
  separate(date, into = c("date", "time"), sep = " ") %>% 
  mutate(
    date = as.Date(date)
  )

all_day_weather_df =
  climate_df %>% filter(time == "23:59:00") %>% 
  distinct(date,daily_weather) %>% 
  filter(!duplicated(date)) %>% 
  separate(daily_weather, into = c("weathertype", "other"), sep = 2) %>%
  mutate(
    weathertype = recode(weathertype,
                      "SN" = "snow",
                      "HZ" = "haze",
                      "RA" = "rain",
                      "FG" = "fog",
                      "BR" = "mist"
                      ),
    weathertype = replace_na(weathertype,"sunny")) %>% 
  select(-other)

collision_all = 
tidy_data %>% 
  group_by(date_complete) %>% 
  summarise(
    collision_event = n(),
    injured_event = sum(persons_injured != 0, na.rm = TRUE),
    killed_event = sum(persons_killed != 0, na.rm = TRUE)
  )

collision_weather = 
bind_cols(all_day_weather_df,collision_all) %>% 
  select(-date_complete)
```

## Weather type
```{r}
weather_type_plot =  
collision_weather %>% 
  group_by(weathertype) %>% 
  summarise(
    count = mean(collision_event)
  ) %>% 
  mutate(
    weathertype = fct_reorder(weathertype,count)
  ) %>% 
  ggplot(aes(x = weathertype, y = count,
             fill = weathertype)) + 
  geom_bar(stat = "identity", width = 0.6) +
  coord_flip()
  
ggplotly(weather_type_plot)
```

## Daily weather
```{r}
daily_weather_df = 
climate_df %>% 
  separate(time, into = c("hour", "minute","second"), sep = ":") %>% 
  separate(date, into = c("year","month", "day"), sep = "-") %>% 
  separate(hourly_precipitation, into = c("prep","other"), sep = 4) %>% 
  select(-minute,-second,-year,-daily_weather,-other) %>% 
  mutate(
    month = as.numeric(month),
    day = as.numeric(day),
    hour = as.numeric(hour),
    prep = replace_na(as.numeric(prep),0)
  )

daily_weather_final = daily_weather_df[rep(seq_len(nrow(daily_weather_df)), each = 3), ]

daily_collision_data = 
tidy_data %>% 
  select(month,day,hour,persons_injured,persons_killed) %>% 
  group_by(month,day,hour) %>% 
  summarise(
    accident = n(),
    injured = sum(persons_injured,na.rm = TRUE),
    killed = sum(persons_killed,na.rm = TRUE)
  ) %>% 
  pivot_longer(
    accident:killed,
    names_to = "events",
    values_to = "number"
  )

daily_weather_collision = 
  left_join(daily_collision_data, daily_weather_df,
            by = c("month","day","hour")) %>% 
  mutate(
    prep_degree = 
      if_else(prep %in% 0:0.10,"light",
              if_else( prep < 0.3 & prep > 0.11 , "moderate","heavy")),
  prep_degree = factor(prep_degree, levels =  c("heavy","moderate","light")))

daily_weather_plot =
daily_weather_collision %>% 
  ggplot(aes(hour, y = number, color = events)) +
  geom_point(alpha = 0.5) + 
  facet_grid(~prep_degree) +
  labs(
    color = "events"
  )

ggplotly(daily_weather_plot)
```

