---
title: "Final Project"
date: "2019/11/10"
output: html_document
---

```{r}
library(tidyverse)
library(plotly)
library(rvest)
library(readr)
library(viridis)
library(leaflet)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

```{r}
data_2018 = 
  read_csv("./data/2018data.csv") %>% 
  janitor::clean_names() 
newnames = colnames(data_2018) %>% 
  str_replace("number_of_","") 
names(data_2018) = newnames

tidy_data = 
data_2018 %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  separate(time, into = c("hour", "minute"), sep = ":") %>%
  select(-zip_code, -location, -on_street_name, -cross_street_name, -off_street_name,-collision_id,-year) %>% 
  rename("vehicle_type" = "vehicle_type_code_1") %>% 
  mutate( day = as.numeric(day),
          month = as.numeric(month),
          hour = as.numeric(hour),
          minute = as.numeric(minute),
          latitude = replace_na(latitude,0),
          vehicle_type = str_to_lower(vehicle_type)
  ) %>%
  filter( latitude != 0)
  
  
```


### Vehicle type
```{r}
vehicle_type_data = 
  tidy_data %>% 
  mutate(
    vehicle_type = replace(vehicle_type,str_detect(vehicle_type,"truck"),"truck"),
    vehicle_type = replace(vehicle_type,str_detect(vehicle_type,"sport utility"),"sport utility vehicle")
    ) %>% 
  filter( vehicle_type %in% c("taxi","passenger vehicle","truck","sport utility vehicle")) %>% 
  group_by(vehicle_type,hour) %>% 
  summarize(
    n = n()
  )

vehicle_type_data %>% 
  plot_ly(
    x = ~hour, y = ~n, color = ~vehicle_type, type = "scatter", mode = "line") %>% 
  layout(
    title = "Collisions of Day for Different Vehicles",
    xaxis = list(title = "Hour of Day"),
    yaxis = list(title = "Collisions")
    )
  
```


### Top 8 Collision Reasons
```{r}
reason_data = 
  tidy_data %>%
  group_by(contributing_factor_vehicle_1) %>%
  summarize(n = n()) %>% 
  arrange(desc(n)) %>% 
  head(10)

reason_data %>% 
  plot_ly(x = ~reorder(contributing_factor_vehicle_1,desc(n)), y = ~n, color = ~contributing_factor_vehicle_1 ,type = "bar") %>% 
  layout(
    title = "The Number of Items Ordered in Each Aisle",
    xaxis = list(title = "Different Reasons"),
    yaxis = list(title = "Count")
    )

```

## Mapping

```{r}
data_2018 = tidy_data

data_2018 = rename(data_2018, long = latitude, lat = longitude)

pal <- colorNumeric(
  palette = "viridis",
  domain = data_2018$persons_injured)



data_2018 %>% 
  filter(!(lat < "-70" | lat >= "-75")) %>% 
  filter(persons_injured > 2) %>% 
    mutate(
    label = str_c("<b>vehicle type: ", vehicle_type, "</b><br>Month: ", month , sep = "") ) %>% 
  sample_n(2000) %>% 
  leaflet() %>% 
  addTiles() %>%
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addLegend("bottomright", pal = pal, values = ~persons_injured,
    title = "Persons Injured",
    opacity = 1
  ) %>% 
  addCircleMarkers(
    ~lat, ~long,
    color = ~pal(persons_injured),
    radius = 0.5,
    popup = ~ label) 

data_2018 %>% 
  group_by(borough) %>% 
  summarise(n())

```

##proportion of accident and injured people by borough, by hour

```{r cars}
data_2018 <- 
  read_csv("./data/2018data.csv") %>% 
  janitor::clean_names() 
newnames = colnames(data_2018) %>% 
  str_replace("number_of_","") 
names(data_2018) = newnames

tidy_data = 
data_2018 %>% 
  separate(date, into = c("day", "month", "year"), sep="/") %>% 
  separate(time, into = c("hour", "minute"), sep=":") %>%
  select(-zip_code, -location, -on_street_name, -cross_street_name, -off_street_name,-year, -collision_id) %>% 
  mutate( day = as.numeric(day),
          month = as.numeric(month),
          hour = as.numeric(hour),
          minute = as.numeric(minute),
          latitude = replace_na(latitude,0)
  ) %>% 
  filter( latitude != 0)
```
```{r}
boro_plot = tidy_data %>% 
  drop_na(borough, persons_injured, persons_killed) %>%
  group_by(borough) %>%
  summarize(accident = n(),
            injured = sum(persons_injured)) %>%
  mutate(prop_accident = round(accident/sum(accident),4),
         prop_injured = round(injured/sum(injured),4)) %>%
  mutate(text_label1 = str_c("accident: ", accident), 
         text_label2 = str_c("injured: ", injured)) %>% 
  plot_ly(x = ~borough, y = ~prop_accident, type = 'bar', name = 'accident', text = ~text_label1, alpha = 0.8) %>%
  add_trace(x = ~borough, y = ~prop_injured, type = 'bar', name = 'injury', text = ~text_label2, alpha = 0.8) %>%
  layout(title="proportion of accident and injured people by borough", xaxis = list(title = 'borough'), barmode = 'group', legend = list(orientation = "h", xanchor = "center", x = 0.5, y = -0.2)) 
boro_plot
```
Number of accident and injured people in Brooklyn is greatest comparing with other boroughs. Queen is second. Staten Island have fewest accident and injured people. Manhattan and Bronx is in the middle.

```{r}
hour_table = tidy_data %>% 
  drop_na(hour, persons_injured, persons_killed) %>%
  group_by(hour) %>%
  summarize(accident = n(),
            injured = sum(persons_injured)) 
hour_plot = hour_table %>%
  mutate(prop_accident = round(accident/sum(accident),4),
         prop_injured = round(injured/sum(injured),4)) %>%
  mutate(text_label1 = str_c("accident: ", accident), 
         text_label2 = str_c("injured: ", injured)) %>% 
  plot_ly(x = ~hour, y = ~prop_accident, type = 'scatter', mode = 'lines', name = 'accident', text = ~text_label1, alpha = 0.8) %>%
  add_trace(x = ~hour, y = ~prop_injured, type = 'scatter', mode = 'lines', name = 'injury', text = ~text_label2, alpha = 0.8) %>%
  layout(title="proportion of accident and injured people by hour", xaxis = list(autotick = FALSE, ticks = "outside", tick0 = 0, dtick = 1, title = 'Hour'), yaxis = list(autotick = FALSE, ticks = "outside", tick0 = 0, dtick = 0.01, title = 'proportion'), legend = list(orientation = "h", xanchor = "center", x = 0.5, y = -0.2)) 

hour_plot %>%
  layout(shapes = list(type = "rect", fillcolor = "pink", line = list(color = "pink"), opacity = 0.3, x0 = 8, x1 = 19, xref = "x", y0 = 0, y1 = 0.08, yref = "y"))
```
Frequency of accident and injured people in the period of 8 am-19 pm is higher than other period of time. Started from 3 am, number of accident and injured people increase and reach a small peak at 8am. At 5pm, it reaches a big peak. then it started to decrease.

