---
title: "Holidays"
output: 
  html_document: 
    code_folding: hide
    toc: true
    toc_float: true
---

```{r include=FALSE}
library(tidyverse)
library(plotly)
library(rvest)
library(readr)
library(viridis)
library(leaflet)
library(patchwork)
library(RColorBrewer)
library(corrplot)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom",plot.title = element_text(face = "bold")))

```

```{r include=FALSE}
data_2018 = 
  read_csv("./data/2018data.csv") %>% 
  janitor::clean_names() 
newnames = colnames(data_2018) %>% 
  str_replace("number_of_","") 
names(data_2018) = newnames

tidy_data = 
data_2018 %>% 
  mutate(
    date_complete = date 
  ) %>% 
  separate(date, into = c("month", "day", "year"), sep = "/") %>% 
  separate(time, into = c("hour", "minute"), sep = ":") %>%
  mutate(
    date_paste = as.Date(paste("2018",month,day,sep = "-"))
  ) %>% 
  select(-zip_code, -location, -on_street_name, -cross_street_name, -off_street_name,-collision_id,-year) %>% 
  rename("vehicle_type" = "vehicle_type_code_1",
         "cause" = "contributing_factor_vehicle_1") %>% 
  mutate( day = as.numeric(day),
          month = as.numeric(month),
          hour = as.numeric(hour),
          minute = as.numeric(minute),
          latitude = replace_na(latitude,0),
          vehicle_type = str_to_lower(vehicle_type)
  ) %>%
  filter( latitude != 0) %>% 
  drop_na(borough)
```


```{r,include=FALSE}
data_2018 = tidy_data

holiday_data = 
  data_2018 %>%
  select(date_complete, everything()) %>%
  mutate(holiday = "No Holiday") %>% 
  select(date_complete, holiday, everything()) %>% 
  mutate(holiday = replace(holiday, date_complete == "01/01/2018", "New Year's Day"), 
         holiday = replace(holiday, date_complete == "01/15/2018", "Martin Luther King, Jr. Day"),
         holiday = replace(holiday, date_complete == "01/13/2018"|date_complete =="01/14/2018", "Martin Luther King Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "02/19/2018", "Presidents Day"),
         holiday = replace(holiday, date_complete == "02/17/2018"|date_complete =="02/18/2018", "Presidents Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "05/28/2018", "Memorial Day"),
         holiday = replace(holiday, date_complete == "05/26/2018"|date_complete =="05/27/2018", "Memorial Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "07/04/2018", "Independence Day"),
         holiday = replace(holiday, date_complete == "07/07/2018"|date_complete =="07/08/2018", "Independence Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "09/03/2018", "Labor Day"),
         holiday = replace(holiday, date_complete == "09/01/2018"|date_complete =="09/02/2018", "Labor Day Weekend (2 days)"),
         holiday = replace(holiday, date_complete == "11/22/2018", "Thanksgiving Day"),
         holiday = replace(holiday, date_complete == "11/23/2018"|date_complete =="11/24/2018"|date_complete =="11/25/2018", "Thanksgiving Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "12/25/2018", "Christmas Day"),
         holiday = replace(holiday, date_complete == "12/26/2018"|date_complete =="12/27/2018", "Christmas Day Weekend (2 days)")
         )
```

```{r, include=FALSE}
holiday_data =
  holiday_data %>% 
  drop_na(persons_injured)

holiday_data %>% 
  group_by(holiday) %>% 
  summarise(total_injured = sum(persons_injured))
```

```{r, include=FALSE}
bronx = 
  holiday_data %>% 
  filter(borough == "BRONX") %>% 
  group_by(holiday) %>% 
  summarise(bronx_total_injured = sum(persons_injured))
brooklyn = 
  holiday_data %>% 
  filter(borough == "BROOKLYN") %>% 
  group_by(holiday) %>% 
  summarise(brooklyn_total_injured = sum(persons_injured)) %>% 
  select(brooklyn_total_injured)
manhattan = 
  holiday_data %>% 
  filter(borough == "MANHATTAN") %>% 
  group_by(holiday) %>% 
  summarise(manhattan_total_injured = sum(persons_injured))%>% 
  select(manhattan_total_injured)
queens = 
  holiday_data %>% 
  filter(borough == "QUEENS") %>% 
  group_by(holiday) %>% 
  summarise(queens_total_injured = sum(persons_injured))%>% 
  select(queens_total_injured)
staten_island = 
  holiday_data %>% 
  filter(borough == "STATEN ISLAND") %>% 
  group_by(holiday) %>% 
  summarise(staten_island_total_injured = sum(persons_injured))%>% 
  select(staten_island_total_injured)

holiday_borough = 
  cbind(bronx, brooklyn, manhattan, queens, staten_island) %>% 
  mutate(total = bronx_total_injured+brooklyn_total_injured+manhattan_total_injured+    queens_total_injured+staten_island_total_injured) %>% 
  filter(holiday!="No Holiday")
```

```{r, include=FALSE}
holiday_borough_day = 
  holiday_borough %>% 
  filter(holiday =="Christmas Day"|holiday =="Independence Day"|holiday =="Labor Day"|holiday =="Martin Luther King, Jr. Day"|holiday =="Memorial Day"|holiday =="New Year’s Day"|holiday =="Thanksgiving Day"|holiday =="Presidents Day")
```

```{r, include=FALSE}
holiday_borough_day = holiday_borough_day[order(holiday_borough_day$total),]

# On the day of the holiday

plot_day =
pivot_longer(
    holiday_borough_day, 
    bronx_total_injured:staten_island_total_injured,
    names_to = "borough", 
    values_to = "person_injured") %>% 
  mutate(holiday = fct_reorder(holiday, -total)) %>% 
  ggplot(aes(x = holiday, y = person_injured, fill = borough)) +
  geom_bar(stat="identity") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 45)) +
  labs(
    title = "Persons Injured by Car Crashs \n on Federal Holidays (2018)",
    x = "Holiday",
    y = "Number Persons Injured") +
  scale_fill_discrete(name = "Borough", labels = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"))
```

```{r, include=FALSE}
data_2018 = tidy_data

holiday_data_weekend = 
  data_2018 %>%
  select(date_complete, everything()) %>%
  mutate(holiday = "No Holiday") %>% 
  select(date_complete, holiday, everything()) %>% 
  mutate(holiday = replace(holiday, date_complete == "01/15/2018"|date_complete == "01/13/2018"|date_complete =="01/14/2018", "Martin Luther King Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "02/17/2018"|date_complete == "02/19/2018"|date_complete =="02/18/2018", "Presidents Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "05/28/2018"|date_complete == "05/26/2018"|date_complete =="05/27/2018", "Memorial Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "07/04/2018"|date_complete == "07/07/2018"|date_complete =="07/08/2018", "Independence Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "09/03/2018"|date_complete == "09/01/2018"|date_complete =="09/02/2018", "Labor Day Weekend (3 days)"),
         holiday = replace(holiday, date_complete == "11/22/2018"|date_complete == "11/23/2018"|date_complete =="11/24/2018"|date_complete =="11/25/2018", "Thanksgiving Weekend (4 days)"),
         holiday = replace(holiday, date_complete == "12/25/2018"|date_complete == "12/26/2018"|date_complete =="12/27/2018", "Christmas Day Weekend (3 days)")
         )

holiday_data_weekend =
  holiday_data_weekend %>% 
  drop_na(persons_injured)

holiday_data_weekend %>% 
  group_by(holiday) %>% 
  summarise(total_injured = sum(persons_injured))
```

```{r include=FALSE}
bronx = 
  holiday_data_weekend %>% 
  filter(borough == "BRONX") %>% 
  group_by(holiday) %>% 
  summarise(bronx_total_injured = sum(persons_injured))
brooklyn = 
  holiday_data_weekend %>% 
  filter(borough == "BROOKLYN") %>% 
  group_by(holiday) %>% 
  summarise(brooklyn_total_injured = sum(persons_injured)) %>% 
  select(brooklyn_total_injured)
manhattan = 
  holiday_data_weekend %>% 
  filter(borough == "MANHATTAN") %>% 
  group_by(holiday) %>% 
  summarise(manhattan_total_injured = sum(persons_injured))%>% 
  select(manhattan_total_injured)
queens = 
  holiday_data_weekend %>% 
  filter(borough == "QUEENS") %>% 
  group_by(holiday) %>% 
  summarise(queens_total_injured = sum(persons_injured))%>% 
  select(queens_total_injured)
staten_island = 
  holiday_data_weekend %>% 
  filter(borough == "STATEN ISLAND") %>% 
  group_by(holiday) %>% 
  summarise(staten_island_total_injured = sum(persons_injured))%>% 
  select(staten_island_total_injured)

holiday_borough_weekend = 
  cbind(bronx, brooklyn, manhattan, queens, staten_island) %>% 
  mutate(total = bronx_total_injured+brooklyn_total_injured+manhattan_total_injured+    queens_total_injured+staten_island_total_injured) %>% 
  filter(holiday!="No Holiday")
```

```{r}
plot_weekend = 
pivot_longer(
    holiday_borough_weekend, 
    bronx_total_injured:staten_island_total_injured,
    names_to = "borough", 
    values_to = "person_injured") %>% 
  mutate(holiday = fct_reorder(holiday, -total)) %>% 
  ggplot(aes(x = holiday, y = person_injured, fill = borough)) +
  geom_bar(stat="identity") +
  theme(legend.position = "right", axis.text.x = element_text(angle = 45)) +
  labs(
    title = "Persons Injured by Car Crashs on \n Federal Holidays Weekends (2018)",
    x = "Holiday Weekends",
    y = "Number Persons Injured") +
  scale_fill_discrete(name = "Borough", labels = c("Bronx", "Brooklyn", "Manhattan", "Queens", "Staten Island"))

plot_weekend = ggplotly(plot_weekend)
plot_weekend

plot_day = ggplotly(plot_day)
plot_day
```

### Comments: 

Most persons injured happens on the day of Labor Day and Labor Day weekend, while least persons injured in on Thanksgiving Day. However, Thanksgiving weekend is the second most holiday weekends when most persons are injured. It could be explained by that people tend to stay home on the day of thanksgiving, while most people might be out shopping on Thanksgiving weekend, which increases the number of persons injured. 

So try to avoid going out on Labor Day weekend and Thanksgiving week, and be careful if you have to. 

